Supplementary Information

### 2021-06-09 doc/manuscript/ForestGEO_treerings.Rmd
comments from Google Docs-- methods

Data analysis consisted of two main steps: (1) identifying the primary climate drivers (i.e., variables and seasonal windows over which they are most influential on tree growth), and (2) combining these climate drivers, DBH, and year into a multivariate model (Fig. `r schematic`). 
The analysis was run separately for each site (step 1), site-species combination (step 2), and each response variable estimating different measures of tree growth (RW, BAI, or $\Delta$AGB). 
We note that the decision to identify primary climate drivers at the level of site, as opposed to species, was motivated by the expectation that differences in the most influential climate drivers across species in one site would be minimal compared to cross-site differences (Fig. `r SCBI_example`); however, analyses focused on interspecific differences could optimize species-specific climate sensitivity estimates by fitting individually by species.

@@

We next combined the primary climate drivers in temperature and precipitation variable groups (included in all models) and DBH (included in models with DBH and its climate interactions) into linear mixed-effects models [function *lme* in the R package *nlme*, @pinheiro_nlme_2021], with core identity as a random intercept and yeaar as a continuous time covariate for the within-group correlation structure (function *corCAR1*). 
We will refer to this model as a generalized least squares (GLS) model (Fig. `r schematic`).

@@

*lme(log[RW] ~ PET + I(PET^2) + $T_{min}$ + I($T_{min}$^2) + DBH + I(DBH^2)",  random = ~1|coreID, correlation = corCAR1(form=~yeaar|coreID), data = x, na.action = "na.fail", method = "ML")*

where *x* is a complete data set (with no missing values) for one species at one site. The method is set to maximum likelihood (*ML*) during the fixed effect model selection phase, but to restricted maximizing likelihood (*REML*) for parameter estimation with the best model.

>>>>

Data analysis consisted of two main steps: (1) identifying the primary climate drivers (i.e., variables and seasonal windows over which they are most influential on tree growth), and (2) combining these climate drivers, DBH, and year into a multivariate model (Fig. `r schematic`). 
The analysis was run separately for each site (step 1), site-species combination (step 2), and each response variable estimating different measures of tree growth (RW, BAI, or $\Delta$AGB). 
We note that the decision to identify primary climate drivers at the level of site, as opposed to species, was motivated by the expectation that differences in the most influential climate drivers across species in one site would be small compared to cross-site differences (Fig. `r SCBI_example`); however, analyses focused on interspecific differences could optimize species-specific climate sensitivity estimates by fitting individually by species.

@@

We next combined the primary climate drivers in temperature and precipitation variable groups (included in all models) and DBH (included in models with DBH and its climate interactions) into linear mixed-effects models [function *lme* in the R package *nlme*, @pinheiro_nlme_2021], with core identity as a random intercept and year as a continuous time covariate for the within-group correlation structure (function *corCAR1*). 
We will refer to this model as a generalized least squares (GLS) model (Fig. `r schematic`).

@@

*lme(log[RW] ~ PET + I(PET^2) + $T_{min}$ + I($T_{min}$^2) + DBH + I(DBH^2)",  random = ~1|coreID, correlation = corCAR1(form=~year|coreID), data = x, na.action = "na.fail", method = "ML")*

where *x* is a complete data set (with no missing values) for one species at one site. The method is set to maximum likelihood (*ML*) during the fixed effect model selection phase, but to restricted maximum likelihood (*REML*) for parameter estimation with the best model.
